\documentclass[./main.tex]{subfiles}
\begin{document}

\doublespacing
Career trajectories---the pathways individuals take throughout their professional lives---have become increasingly complex in the modern workplace. No longer confined to linear progressions within single organizations, today's career journeys involve lateral moves, industry transitions, skill pivots, and varied advancement patterns. This comprehensive volume presents 23 distinct methodological approaches for analyzing career trajectories, organized within six overarching categories that represent the breadth of analytical frameworks available to researchers, practitioners, and policymakers.

\subsection{Purpose and Organization}\label{purpose-and-organization}
This literature overview serves as a comprehensive resource for
understanding the diverse methodological landscape of career trajectory analysis. Each approach is examined through a consistent analytical framework, enabling readers to compare methodologies across categories while appreciating the unique contributions each makes to our understanding of career development. The 23 approaches are systematically organized into six categories, reflecting different analytical perspectives and disciplinary traditions:

\subsubsection{Descriptive and Classification Methods}\label{descriptive-and-classification-methods}
These approaches describe, classify, or map career paths and states, often using empirical data to identify patterns, archetypes, or categories. This category includes Career Path Mapping, Sequence Analysis/Optimal Matching Analysis, Latent Class Analysis, Job Classification, Historical Trend Analysis, and Demographic Analysis. These methods excel at revealing underlying structures within career data.

\subsubsection{Longitudinal and Cohort
Analysis}\label{longitudinal-and-cohort-analysis}
These approaches track individuals or groups over time to observe changes, transitions, and career development. Methods include Longitudinal Cohort Studies, Retention and Attrition Patterns, and Promotion Board Analysis, focusing on temporal aspects of career progression.

\subsubsection{Network and Relationship
Analysis}\label{network-and-relationship-analysis}

These approaches examine how social structures, networks, and relationships shape career trajectories. Social Network Analysis and Mentorship and Sponsorship Studies emphasize the relational aspects of career advancement.

\subsubsection{Quantitative Performance and Outcome
Analysis}\label{quantitative-performance-and-outcome-analysis}

Measuring and analyzing performance, advancement, and outcomes using quantitative data is the focus of approaches like Quantitative Performance Metrics, Comparative Branch/Service Analysis, Career Field Health Assessment, and Education and Training Impact Assessment.

\subsubsection{Modeling and Simulation}\label{modeling-and-simulation}

These approaches use computational, statistical, or mathematical models to simulate, forecast, or optimize career-related processes, including Agent-Based Modeling, Monte Carlo Simulations, System Dynamics Models, Strategic Workforce Planning Models, Training Pipeline Optimization, and Force Structure Projections.

\subsubsection{Policy and Impact Evaluation}\label{policy-and-impact-evaluation}
These approaches assess the effects of interventions, policies, or changes on career outcomes, incorporating Policy Impact Evaluation and Policy-Testing Simulations.

\subsection{Chapter Structure: The Analytical Framework}\label{chapter-structure-the-analytical framework}
Each of the 23 approaches is examined through a consistent eight-part framework, allowing readers to fully understand each methodology while facilitating meaningful comparisons across approaches. The framework structure used throughout the literature overview is as follows:

\subsubsection{Approach Description \& Goal}\label{approach-description-goal}
Each chapter begins with a concise description of the approach and its primary objectives. For example, in Career Path Mapping, this section explains how the approach identifies transitions between occupational roles and visualizes common advancement patterns. For Latent Class Analysis, it describes how the method identifies unobserved subgroups within career data based on response patterns.

\subsubsection{Critical Variables}\label{critical-variables}
This section details the key variable categories typically used as inputs to the approach. Demographic Analysis might highlight variables such as gender, age cohorts, and educational attainment levels. Retention and Attrition Patterns would emphasize variables like time-to-separation, risk factors for early departure, and organizational tenure metrics.

\subsubsection{Key Overviews}\label{key-overviews}
Each chapter provides paragraph summaries of 2-4 seminal works that introduce the approach. For Sequence Analysis, this might include Abbott's foundational work on social sequences and Optimal Matching techniques. For Strategic Workforce Planning Models, it would cover key frameworks for integrating labor supply forecasting with organizational demands.

\subsubsection{Mathematical Approach}\label{mathematical-approach}
Here, each chapter explains the analytical logic underpinning the method, including relevant formulas. The Social Network Analysis chapter details network density calculations (D = 2E/n(n-1)), centrality measures, and graph theory applications. The Monte Carlo Simulations section explains probability distributions and random sampling techniques used in career trajectory forecasting.

\subsubsection{Example Applications}\label{example-applications}
Each methodology chapter includes paragraph summaries of 2-4 illustrative studies applying the approach to career trajectory analysis. For Job Classification, this might feature research on competency-based frameworks in healthcare settings. For System Dynamics Models, it could highlight studies modeling workforce flows in military organizations.

\subsubsection{Critiques}\label{critiques}
This section presents balanced criticism of each approach, with references to scholarly critiques. For example, Agent-Based Modeling faces challenges in parameter calibration and validation, while Historical Trend Analysis may struggle with applicability during periods of rapid technological change.

\subsubsection{Software}\label{software}
Each chapter describes the primary software tools used for implementing the methodology. For Longitudinal Cohort Studies, this includes R packages for multilevel modeling and Python libraries for handling panel data. For Career Field Health Assessment, it details specialized workforce analytics platforms.

\subsubsection{Example Study Design}\label{example-study-design}
The final section provides a practical study design exemplifying the approach, including Key Variables, Sample \& Data Collection, Analysis Approach, Potential Findings, and Potential Implications. The example study design is purposely geared toward usage and analysis of potential indicators as identified by the UVA Bioinformatics team, summer 2024 (see below). 

\singlespacing
\begin{tabularx}{6.5in}{|l|X|X|X|}

\hline
 &\multicolumn{3}{|l|}{U.S. Army Branch Indicators} 
 \\
\hline
 & Armor & Logistics & Additional Notes \\
\hline
Education &
\begin{itemize}
    \item Commission Source
    \item Elective Training
\end{itemize}
&
\begin{itemize}
    \item Commission Source
    \item Elective Training
    \item Graduate Education
    \item Skill Identifier (SI)
    \item Secondary AOC
\end{itemize}
&
\rr
Listed examples are often competitive and selective that are optional or represent alternative paths for PME \newline
Commission source has significant effect on survival curves of U.S. army officers (Doganca 2006)
\tn
\hline
Experience &
\rr
\begin{itemize}
    \item Key Development (KD)
    \item Broadening Assignments
    \item Joint Service Credit
\end{itemize}
 &
\rr
\begin{itemize}
    \item Key Development (KD)
    \item Broadening Assignments
\end{itemize}
&
\rr
Performance in KD assignment heavily influences potential for promotion \newline
BSB/CSSB S3 position provided to exceptional captains \newline
mGPA is strong determinant as similar to jobs performed as officers (Spain 2020)
\tn
\hline
Assessments &
\rr
\begin{itemize}
\item Officer Evaluation Report (OER) 
\item Battalion Commander Assessment 
\item Commander Assessment Program 
\item Personality Assessment 
\end{itemize}
&
\rr
\begin{itemize}
    \item Battalion Commander Assessment
    \item Commander Assessment Program
    \item BOLC Assessment
    \item CCC Assessments
    \item Officer Evaluation Report (OER)
\end{itemize}
&
JDAL position or JQS points to measure JIIM (Joint Interagency Intergovernmental Multinational Environment) or diversity of experience \\
\hline
Progression &
\rr
\begin{itemize}
    \item Captain (5 years)
    \item Major (10 years)
    \item Lieutenant Colonel (16 years)
    \item Colonel (22 years)
\end{itemize}

&
\rr
\begin{itemize}
    \item Captain (4 years)
    \item Major (10.5 years)
    \item Lieutenant Colonel (16 years)
    \item Colonel (23 years)
\end{itemize}

&
\rr
Promotion is more difficult beyond O-3 (Captain) as target promotion from O-3 to O-4 is \(80\%\) to \(70\%, 50\%, 10\%\) with increasing service duration required (Doganca 2006) 
\tn
\hline
\end{tabularx} \\

\begin{tabularx}{6.5in}{|l|X|X|X|}
\hline
 & \multicolumn{1}{|c|}{Aviation} & \multicolumn{1}{c|}{Cyber} & \multicolumn{1}{c|}{Additional Notes} \\
\hline
Education &
\rr
\begin{itemize}
    \item Senior Service College (SSC)
    \item Pre-Combat Checks (PCC)
    \item Elective Training
\end{itemize}
&
\rr
\begin{itemize}
    \item Senior Service College (SSC)
    \item Elective Training
    \item Commission Courses
\end{itemize}
&
\rr
The options are often competitive and elective training is highly recommended.
\tn
\hline
Experience &
\rr
\begin{itemize}
    \item Key Developments (KD)
    \item Duty Assignments
\end{itemize}
&
\rr
\begin{itemize}
    \item Key Developments (KD)
    \item Duty Assignments
\end{itemize}
&
\rr
Emphasizing harder assignments and work done in the assignments over time doing them \\
For Aviation the choice of KD is a lot smaller than the choices given to Cyber
\tn
\hline
Assessments &
 & 
 \rr
\begin{itemize}
    \item Cyber Aptitude and Talent Assessment (CATA)
    \item Cyber Test (CT)
    \item Cyber Assessment and Selection Program (CASP)
\end{itemize}
&
\rr
Didn't find this information in either of the docs that were sent but did some research on my own and found this information (Morris 2020) \\
Was unable to find anything specific for Aviation
\tn
\hline
Progression &
\rr
\begin{itemize}
    \item Captain (4 years) 
    \item Major (10 years) 
    \item Lieutenant Colonel (16 years) 
    \item Colonel (20 years)
\end{itemize}
&
\rr
\begin{itemize}
    \item Captain (4 years) 
    \item Major (11 years) 
    \item Lieutenant Colonel (17 years) 
    \item Colonel (22 years)
\end{itemize}
&
\tn
\hline
\end{tabularx}



\begin{tabularx}{6.5in}{|l|X|}
\hline
\multicolumn{2}{|l|}{Additional Indicators} 
\\
\hline
 & Additional Notes \\
\hline
\rr
Armed Services Vocational Aptitude Battery (ASVAB) & Major cognitive abilities measures are useful in predicting job performance (Ree 1994) \\
\hline
Age & Higher age has correlation with survival curve of U.S. army officer (Doganca 2006) \\
\hline
Race & African Americans have correlation with survival curve of U.S. army officer (Doganca 2006) \\
\hline
Gender & Females has correlation with survival curve of U.S. army officer (Doganca 2006) \\
\hline
Marital Status & Spousal dependents have correlation with survival curve of U.S. army officer (Doganca 2006) \\
\hline 
\end{tabularx}



\begin{tabularx}{6.5in}{|l|X|}
\hline
\multicolumn{2}{|l|}{Non-Cognitive Attributes} 
\\
\hline
 & Additional Notes \\
\hline
Hardiness-control & Significant correlation with military performance both at West Point and as an officer as measured by Reception Day battery of tests (Bartone 2013) \\
\hline
Hardiness-commitment & Significant correlation with military performance both at West Point and as an officer (Bartone 2013) \\
\hline
Hardiness-challenge & Do not perform well in conventional structured military and leadership tasks (Bartone 2013) \\
\hline
Motivation & Motivation was strongly associated with job performance (Doganca 2006) \\
\hline
Complex problem solving & Complex problem solving was consistent predictor of continuance and advancement (Zaccaro 2012) \\
\hline
Creative thinking & Creative thinking was consistent predictor of continuance and advancement (Zaccaro 2012) \\
\hline
Responsibility & Responsibility was consistent predictor of continuance and advancement (Zaccaro 2012) \\
\hline
\end{tabularx}

\doublespacing

\subsection{The Value of a Unified
Framework}\label{the-value-of-a-unified-framework}

By presenting each methodology through this consistent lens, this volume
offers several advantages:

First, it enables meaningful comparisons across approaches. Readers can
easily identify when Quantitative Performance Metrics might be more
appropriate than Comparative Branch/Service Analysis for a specific
research question, or understand how Social Network Analysis complements
Mentorship and Sponsorship Studies.

Second, it bridges disciplinary boundaries. Methods originating in
sociology (Sequence Analysis), psychology (Latent Class Analysis),
economics (Policy Impact Evaluation), and computer science (Agent-Based
Modeling) are presented in a unified format accessible to scholars and
practitioners from diverse backgrounds.

Third, it facilitates methodological innovation. By understanding the
strengths, limitations, and underlying mathematics of multiple
approaches, researchers can develop hybrid methodologies that address
complex career questions from multiple angles.

Fourth, it supports practical application. Human resource professionals,
career counselors, and workforce planners can select appropriate
methodologies based on their specific contexts, available data, and
research objectives.

This volume represents the first comprehensive effort to systematically
catalog and explain the diverse methodological landscape of career
trajectory analysis as it applies to the careers of military officers.
Whether you are a researcher seeking methodological rigor, a
practitioner aiming to implement evidence-based career development
programs, or a policymaker evaluating workforce interventions, this
volume provides the analytical tools needed to understand the
increasingly complex pathways of modern careers.




\end{document}
